---
title: "ggplot2"
author: "Padma Tanikella"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this activity, we will use the CPS85 data from the mosaicData package.
```{r}
library(mosaicData)
head(CPS85)
```

#### Question 1
Provide a brief description of the dataset.



#### Question 2
Last time, we saw how `facet_wrap` works. (Refer to lecture ) Today, we will try to understand `facet_grid`.


* `facet_wrap()` just makes a seperate plot for each level of the categorical variable
  * `facet_wrap( ~ categoricalVariable)`
  
* `facet_grid()` allows control of row & column facets
* `facet_grid()` syntax:
  * row and column facets: `facet_grid(rows ~ cols)`
  * row facets only: `facet_grid( rows ~ . )` (note the required “`.`”)
  * column facets only: `facet_grid( ~ cols)` (no “`.`” this time)

Using the CPS85 data, provide a useful visual summary using `facet_grid`. 


#### Question 3
```{r}
## code 1
CPS85 %>% 
  ggplot() +
  geom_density(aes(x = wage, color = sex), alpha = 0.4)+
  facet_grid( ~ married) +
  xlim(0,30) 


## code 2
CPS85 %>% 
  ggplot() +
  geom_density(aes(x = wage, fill = sex), alpha = 0.4)+
  facet_grid( ~ married) +
  xlim(0,30)


## code 3
CPS85 %>% 
  ggplot() +
  geom_density(aes(x = wage, fill = sex, color = sex), alpha = 0.4)+
  facet_grid( ~ married) +
  xlim(0,30)
```


1. Notice that code 1 and code 2 are the same except for `color = sex` in code 1 and `fill = sex` in code 2. Look at the outputs of code 1 and code 2 and explain the different between the aesthetics `fill` and `color`.

#### Question 4
This website provides basic information on how to construct a histogram using `ggplot2`: https://r-charts.com/distribution/histogram-binwidth-ggplot2/

Using the CPS85 data, provide a useful visual summary using `geom_histogram`. Number of bins is a subjective choice. Look at how the histogram's appearance changes as the number of bins is changed. 










