---
title: "L07 - Data Wrangling"
author: 
  - Presenter- Padma Tanikella 
  - Content Credit- Dr. Matthew Beckman
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today's class will involve an in class activity where you practice data wrangling. 

## Agenda

- Introduce some software and commands that ...
    - make it easy to access data tables and see how they are structured
        - For example: `data()`, `View()`, `help()`, 
        - (more coming in Chapter 10)
    - learn about data verbs 
    
    
## Three Important Concepts 

1. Data can be usefully organized into tables with "cases" and "variables."  
    - In "tidy data" every case is the same sort of thing (e.g. a person, a car, a year, a country in a year) 
    - We sometimes even modify data in order to change what the cases represent in order to better represent a point.

2. Data graphics and "glyph-ready" data
    - each case corresponds to a "glyph" (mark) on the graph
    - each variable to a graphical attribute of that glyph such as x- or y-position, color, size, length, shape, etc. 
    - same is true for more technical tools (e.g., models, predictions, etc.) 

3. When data are not yet in glyph-ready form, you can transform (i.e. wrangle) them into glyph-ready form.  
    - Such transformations are accomplished by performing one or more of a small set of basic operations on data tables
    - This is the work of data "verbs" 
    
    
## Learning about the raw data

There are lots of ways to load data into your environment

- Most real data sources will require you to 
    - read a file (e.g., CSV) 
    - query a database (e.g., SQL)
    - configure an API
    - scrape from the web
- For convenience, many STAT 184 data sets are accessed from R packages or CSV files
- When acquiring data, it's very important to pause and think about data provenance/origins
    - What might be useful to learn?
    - How is this accomplished?
    - Why does it matter?
    

## Recall: Key goals of a careful Exploratory Data Analysis?

1. **Examine the data source:** variable types, coding, missingness, summary statistics/plots, who/what/when/where/why/how data were collected
2. **Discover features that influence may modeling decisions:** investigate potential outliers, consideration for recoding variables (e.g., numeric data that's functionally dichotomous), evaluate correlation structure (e.g., autocorrelation, hierarchy, spatial/temporal proximity)
3. **Address research questions:** build intuition and note preliminary observations/conclusions related to each research question.  Also, note observations that prompt you to refine your research questions or add new questions to investigate

## Guided practice 

- `Minneapolis2013` data set in the `dcData` package
  - To do this, we need to download the package from GitHub

```{r}

# Install the package from GitHub
# The very first time you run this, uncomment the 4 lines below

# install.packages("devtools")
# library(devtools)
# library(usethis)
# install_github("mdbeckman/dcData")
library(dcData)
data("Minneapolis2013", package = "dcData")


```


### Q1 : Looking at the dataset
Answer the following questions about the dataset:

1. What is the setting for the data?  
    - What are they about?
    - Who collected them?
    - Why were they collected?
    - etc
2. How many cases are there?
3. In your own words, what kind of thing do the cases represent? 
4. How many variables are there?  What are their names?
5. Pick out three of the variables and say whether
    - the variable is quantitative or categorical
    - if categorical (R calls this a "factor"), what are some levels of the variable
    - if quantitative, what are the units of measurement of the variable.

Click [here](https://en.wikipedia.org/wiki/Instant-runoff_voting) to learn about rank choice voting (also called instant run off voting). 



### Question 2
Create a table called `VoteResults` recording the **first** choice of each of 80101 individual precincts in the 2013 mayoral election. 

```{r}
# Hint: use tidyverse functions

VoteResults <-

```


### Question 3
Based on what we learnt about the grammar of `ggplot2` graphics, construct a _sorted_ bar chart of the `VoteResults` data. Label the x-axis and y-axis appropriately.

```{r}
# Hint: use geom_bar() to construct bar charts
# look at how the reorder() function works

```


### Question 4
Run the following code:
```{r}
VoteResults %>% 
  arrange( desc(votes) ) %>%
  head()   
```

Compare the `Minneapolis2013` data table and the wrangled data table printed above.

1. Do they have the same number of cases?
2. Do the cases in the two tables represent the same sort of thing?
3. Do the two tables have any variable(s) in common?
4. How are the two tables are related to one another?


## Why we wrangle

Data wrangling **prepares** the data for analysis. 

- convert to tidy form for computing
- prepare glyph-ready data for visualization
- prepare data for modeling (e.g., exploratory, inferential, predictive)


## Different types of functions


- **Reduction functions**
    - inputs are **variables** 
    - results are **scalar**
    - examples: `sum()`, `mean()`, `sd()`, `n()`, `n_distinct()`
- **Transformation functions**
    - inputs are **variables**; 
    - results are **variable**
    - examples: `mutate`, `rename`
      - utilizes functions/operators such as +, -, *, /, ^, logicals, `round()`, `log()` 
- **Data verbs**
    - inputs are **data frames**
    - results are **data frames**
    - examples: `summarise()`, `group_by()`



Let's use some other reduction functions, transformation functions, and data verbs with the some NFL data


### Question 5
What is the dataset describing? Use R commands to get a preliminary understanding of the dataset at hand.
```{r}
dat.football <- read_tsv(file = "https://raw.githubusercontent.com/ada-lovecraft/ProcessingSketches/master/Bits%20and%20Pieces/Football_Stuff/data/nfl-salaries.tsv")




```


### Question 6
Use `ggplot2` graphics to construct a graphical summary of the dataset. What information does the graphical summary tell about the dataset.
```{r}

```


### Question 7
What is the highest salary and which player(s) earn that amount?

```{r}

```


### Question 8
What is the team with the highest paid roster, and what was their total pay? 
What is the team with the lowest paid roster, and what was their total pay?

Note: Paid roster is the total salary of all the players in the team.

```{r}

```







