---
title: "L01-Basic R class notes"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Basic Terminology
* Difference between R and RStudio
* Show the components of RStudio - editor, console, environment, final pane has help, output viewer, packages
* $>$ is called prompt. It means that R is ready to take your command(s). 
* Commands are "executed".
* Tell about R sessions.
* Marin stat lectures youtube channel.

# Basic R computions

```{r}
## do these in console and in R script

## mention about writing "neat code"
5 + 6

# show how environment stores a variable called total
## explain R is case sensitive
total <- 11

## similarly we have sin(), cos(), log(), pi
log(12); sin(1); pi; exp(0)

## -Inf and +Inf exist as well
1/0

```


```{r}

# data in base-R
data("iris")
View(iris)
head(iris)
str(iris) # structure of iris data

dim(iris)
mean(iris$Sepal.Length)
sd(iris$Sepal.Width)
```

```{r}
help("plot")
plot(iris$Sepal.Length, iris$Sepal.Width, xlab = "Sepal length", ylab = "Sepal width")
```

# Packages and libraries
* install directly on Rstudio
* using `install.packages()`
* CRAN, bioconductor
* `library()` and talk about "loading" packages.

# Vectors and Matrices

* creating vectors using `c()`
* `seq()`, `t()`, `length()`, `rep()`
* R performs element wise computation
* try computing distance between vectors x and y
* `matrix()`
* For a matrix A, `A*A` vs `A%*%A`
* `diag()`








